Sequential Decision-Making under Uncertainty with Big Data

### 2023-04-06 paper/main.tex

\textbf{Laplace approximation}
The Laplace approximation \todo{add citation of original paper} constructs a Gaussian approximation of the weight-space posterior $p(\transitionParams \mid \dataset)$
by using a second-order Taylor expansion of $\mathcal{L}$ around $\transitionParams^{*}$.
The posterior approximation is given by,
\begin{align} \label{eq-laplace-approx-weight-space}
  p(\transitionParams \mid \dataset) \approx \mathcal{N} \left( \transitionParams \mid \transitionParams^{*} , \bm\Sigma_{\transitionParams^{*}} \right)
  \quad \text{with} \quad \bm\Sigma_{\transitionParams} =
 - \nabla_{\transitionParams \transitionParams}^{2} \mathcal{L} ( \dataset ; \transitionParams)|_{\transitionParams=\transitionParams^{*}}
\end{align}
That is, it sets the posterior precision to the Hessian of the loss at the optimal parameters $\transitionParams^{*}$.
It is worth noting that computing the Hessian of the loss can be computationally intractable for large networks.
The prior terms are usually trivial, so we will focus on the likelihood here.
The Jacobian and Hessian of the log likelihood can be expressed per data point,
\begin{align} \label{eq-jac}
  \nabla_{\transitionParams} \log p(\mathbf{y} \mid \mathbf{f}(\mathbf{x}; \transitionParams)) &= \mathbf{J}(\mathbf{x})^{T} \mathbf{r}(\mathbf{y} ; \mathbf{f}) \\
  \nabla^{2}_{\transitionParams\transitionParams} \log p(\mathbf{y} \mid \mathbf{f}(\mathbf{x}; \transitionParams)) &= \mathbf{H}(\mathbf{x})^{T}
  \mathbf{r}(\mathbf{y};\mathbf{f}) - \mathbf{J}(\mathbf{x})^{T} \bm\Lambda(\mathbf{y};\mathbf{f}) \mathbf{J}(\mathbf{x}),
\label{eq-hess}
\end{align}
through the Jacobian $\mathbf{J} \in \R^{C \times P}$ and Hessian $\mathbf{H} \in \R^{C \times P \times P}$ of the feature extractor $\mathbf{f}(\mathbf{x}; \bm\transitionParams)$,
\begin{align} \label{eq-jac}
[\mathbf{J}(\cdot)]_{ci} = \odv{f_{c}(\cdot ; \transitionParams)}{\transitionParams_{i}}_{\transitionParams=\transitionParams^{*}} \qquad
[\mathbf{H}(\cdot)]_{cij} = \odv{f_{c}(\cdot ; \transitionParams)}{\transitionParams_{i}\transitionParams_{j}}_{\transitionParams=\transitionParams^{*}}
\end{align}
where
$\mathbf{r}(\mathbf{y}; \mathbf{f}) = \nabla_{\mathbf{f}} \log p(\mathbf{y} \mid \mathbf{f})$ can be interpreted as a residual and
$\bm\Lambda(\mathbf{y} ; \mathbf{f}) &= \nabla^{2}_{\mathbf{f} \mathbf{f}} \log p(\mathbf{y} \mid \mathbf{f})$
as per-input noise.
Many approximations exist and can be used alongsied our method.
We refer the reader to \cite{daxbergerLaplace2021} for more details.
See \cref{sec-laplace-approximation} for further details on the Laplace approximation.

>>>>

\textbf{Laplace approximation}
The Laplace approximation \todo{add citation of original paper} constructs a Gaussian approximation of the weight-space posterior $p(\transitionParams \mid \dataset)$
by using a second-order Taylor expansion of $\mathcal{L}$ around $\transitionParams^{*}$.
The posterior approximation is given by,
\begin{align} \label{eq-laplace-approx-weight-space}
  p(\transitionParams \mid \dataset) \approx \mathcal{N} \left( \transitionParams \mid \transitionParams^{*} , \bm\Sigma_{\transitionParams^{*}} \right)
  \quad \text{with} \quad \bm\Sigma_{\transitionParams} =
 - \nabla_{\transitionParams \transitionParams}^{2} \mathcal{L} ( \dataset ; \transitionParams)|_{\transitionParams=\transitionParams^{*}}
\end{align}
That is, it sets the posterior precision to the Hessian of the loss at the optimal parameters $\transitionParams^{*}$.
It is worth noting that computing the Hessian of the loss can be computationally intractable for large networks.
The prior terms are usually trivial, so we will focus on the likelihood here.
The Jacobian and Hessian of the log likelihood can be expressed per data point,
\begin{align} \label{eq-jac}
  \nabla_{\transitionParams} \log p(\mathbf{y} \mid \mathbf{f}(\mathbf{x}; \transitionParams)) &= \mathbf{J}(\mathbf{x})^{T} \mathbf{r}(\mathbf{y} ; \mathbf{f}) \\
  \nabla^{2}_{\transitionParams\transitionParams} \log p(\mathbf{y} \mid \mathbf{f}(\mathbf{x}; \transitionParams)) &= \mathbf{H}(\mathbf{x})^{T}
  \mathbf{r}(\mathbf{y};\mathbf{f}) - \mathbf{J}(\mathbf{x})^{T} \bm\Lambda(\mathbf{y};\mathbf{f}) \mathbf{J}(\mathbf{x}),
\label{eq-hess}
\end{align}
through the Jacobian $\mathbf{J} \in \R^{C \times P}$ and Hessian $\mathbf{H} \in \R^{C \times P \times P}$ of the feature extractor $\mathbf{f}(\mathbf{x}; \bm\transitionParams)$,
\begin{align} \label{eq-jac}
[\mathbf{J}(\cdot)]_{ci} = \odv{f_{c}(\cdot ; \transitionParams)}{\transitionParams_{i}}_{\transitionParams=\transitionParams^{*}} \qquad
[\mathbf{H}(\cdot)]_{cij} = \odv{f_{c}(\cdot ; \transitionParams)}{\transitionParams_{i}\transitionParams_{j}}_{\transitionParams=\transitionParams^{*}}
\end{align}
where
$\mathbf{r}(\mathbf{y}; \mathbf{f}) = \nabla_{\mathbf{f}} \log p(\mathbf{y} \mid \mathbf{f})$ can be interpreted as a residual and
$\bm\Lambda(\mathbf{y} ; \mathbf{f}) = \nabla^{2}_{\mathbf{f} \mathbf{f}} \log p(\mathbf{y} \mid \mathbf{f})$
as per-input noise.
Many approximations exist and can be used alongsied our method.
We refer the reader to \cite{daxbergerLaplace2021} for more details.
See \cref{sec-laplace-approximation} for further details on the Laplace approximation.
